### 2016-05-29 paper/put-together.tex
added latexes

\input{fig_regex-definition}
\input{fig_regex-semantics}
However, if we are to use the types to direct, and the types are
regular expressions, we would like to allow for users to easily engineer
regular expressions.
For this we have created a language of regular expressions with a context,
with syntax defined in Figure~\ref{fig:regex-definition}, and semantics defined
in Figure~\ref{fig:regex-semantics}.  In this, we only allow the context to
have finite size.
We show that adding in a context doesn't increase expressibility.
\begin{restatable}[Completeness of the Regular Expressions Without Variables]{theorem}{ecrc}
\label{thm:empty-context-regex-completeness}
For all contexts \RegexContext{} and regular expressions \Regex{}, there exists a regular expression \Regex{}',
such that \LanguageOf{\RegexContext{}}{\Regex{}}=\LanguageOf{\emptyset}{\Regex{}'}.
\end{restatable}
\begin{proof}
By induction on the size of the context and the regular expression
\end{proof}

@@

Interestingly, the use of lens contexts and regular expression contexts
is not needed.
This can be formalized in the following theorem.
\begin{restatable}[Completeness of Bijective Lenses Without Variables]{theorem}{eclc}
\label{thm:empty-context-lens-completeness}
If $\FullContext \vdash \Lens : \Regex \Leftrightarrow \RegexAlt \HasSemantics \PutRight,\PutLeft$,
then there exists a $\Lens'$, $\Regex'$, $\RegexAlt'$, such that
\LanguageOf{\RegexContext}{\Regex}=\LanguageOf{\emptyset}{\Regex'},
\LanguageOf{\RegexContext}{\RegexAlt}=\LanguageOf{\emptyset}{\RegexAlt'}, and
$\emptyset,\emptyset \DerivesLens \Lens' \OfType \MapsBetweenTypeOf{\Regex'}{\RegexAlt'} \HasSemantics \PutRight,\PutLeft$.
\end{restatable}
\begin{proof}
Through induction on typing derivations.
\end{proof}

@@

\input{fig_dnf-lens-syntax}
\input{fig_dnf-lens-alternate-alternate-semantics}

Because of this, to synthesize lenses, we instead create a new, equivalent language,
who handles many of the equivalences of regular expressions through exploiting a normal form.
We have created a language which can handle many of the equivalences of regular
expressions through its normal form, and adds additional structure to the search
space, constraining it and minimizing the traversal through equivalent elements.
We call this language, the language of DNF lenses, due to the fact these lenses are
put into a form similar to the disjunctive normal form of logic.
The syntax for these lenses is provided in Figure~\ref{fig:dnf-lens-syntax}.
The typing and semantics for these lenses is provided in Figure~\ref{fig:dnf-lens-alternate-alternate-semantics}.

DNF lenses handle the equivalences of concatenation associativity,
or associativity, or commutativity, and and-or distributivity.
Furthermore, they do not require concatenation to be complete,
though we do allow for concatenation of lenses in the lens context.
While these are not needed, this is under the assumption that if lenses are present
in the context, they are likely to be used.
Because of this, the choices needed to be done by the synthesis algorithm
are fairly minimal.  There are only
a few types of equivalences related to stars and use of regular expressions
in the context that need to be explored for completeness.
Furthermore, there is the difficulty of determining which permutations create
well typed regular expressions, but our implentation reduces the problem of
finding a valid permutation, to ordering of terms under a specific order.

>>>>

\input{fig_regex-syntax}
\input{fig_regex-semantics}
However, if we are to use the types to direct, and the types are
regular expressions, we would like to allow for users to easily engineer
regular expressions.
For this we have created a language of regular expressions with a context,
with syntax defined in Figure~\ref{fig:regex-syntax}, and semantics defined
in Figure~\ref{fig:regex-semantics}.  In this, we only allow the context to
have finite size.
We show that adding in a context doesn't increase expressibility.
\begin{restatable}[Completeness of the Regular Expressions Without Variables]{theorem}{ecrc}
\label{thm:empty-context-regex-completeness}
For all contexts \RegexContext{} and regular expressions \Regex{}, there exists a regular expression \Regex{}',
such that \LanguageOf{\RegexContext{}}{\Regex{}}=\LanguageOf{\emptyset}{\Regex{}'}.
\end{restatable}
\begin{proof}
By induction on the size of the context and the regular expression
\end{proof}

@@

Interestingly, the use of lens contexts and regular expression contexts
is not needed.
This can be formalized in the following theorem.
\begin{restatable}[Completeness of Bijective Lenses Without Variables]{theorem}{eclc}
\label{thm:empty-context-lens-completeness}
If $\FullContext \vdash \Lens : \Regex \Leftrightarrow \RegexAlt \HasSemantics \PutRight,\PutLeft$,
then there exists a $\Lens'$, $\Regex'$, $\RegexAlt'$, such that
\LanguageOf{\RegexContext}{\Regex}=\LanguageOf{\emptyset}{\Regex'},
\LanguageOf{\RegexContext}{\RegexAlt}=\LanguageOf{\emptyset}{\RegexAlt'}, and
$\emptyset,\emptyset \DerivesLens \Lens' \OfType \MapsBetweenTypeOf{\Regex'}{\RegexAlt'} \HasSemantics \PutRight,\PutLeft$.
\end{restatable}
\begin{proof}
Through induction on typing derivations.
\end{proof}
However, despite the fact these variables are not needed, they are useful for nice
separation of concerns.
Lenses in the context are likely to be used, much like when programming, one
doesn't import libraries for no reason.

@@

\input{fig_dnf-regex-syntax}
\input{fig_dnf-regex-semantics}
\input{fig_dnf-regex-functions}
\input{fig_dnf-regex-rewrites}
Because of this, to synthesize lenses, we instead create a new, equivalent language,
who handles many of the equivalences of regular expressions through exploiting a normal form.
We define this normal form of regular expression DNF Regular expression, with syntax
defined as in Figure~\ref{fig:dnf-regex-syntax}, and with semantics defined as
in Figure~\ref{fig:dnf-regex-semantics}.  These semantics rely on some functions
defined in Figure~\ref{fig:dnf-regex-functions},
and some rewrite rules defined in Figure~\ref{fig:dnf-regex-rewrites}.
This normal form is called a DNF regular expression, as it is similar to the
sentences in logical disjunctive normal form.
These DNF regular expression expresses exactly the same amount as standard regular
expressions.
\begin{restatable}[Completeness of DNF Regexs]{theorem}{dnfrc}
\label{thm:completeness-dnf-lenses}
There exists a function $\ToDNFRegex\OfType\ArrowTypeOf{\mathit{Regex}}{\mathit{DNF Regex}}$,
such that, for all $\RegexContext$, $\LanguageOf{\RegexContext}{\Regex}=\LanguageOf{\RegexContext}{\ToDNFRegex(\Regex)}$.
\end{restatable}

\begin{restatable}[Soundness of DNF Regexs]{theorem}{dnfrs}
\label{thm:soundness-dnf-lenses}
The function \ToDNFRegex{} has a right inverse \FromDNFRegex{}.
\end{restatable}

\input{fig_dnf-lens-syntax}
\input{fig_dnf-lens-alternate-alternate-semantics}
We have created a language which can handle many of the equivalences of regular
expressions through operation on dnf regular expressions,
and adds additional structure to the search
space, constraining it and minimizing the traversal through equivalent elements.
We call this language, the language of DNF lenses, due to the fact the lenses
operate on DNF regular expressions.
The syntax for these lenses is provided in Figure~\ref{fig:dnf-lens-syntax}.
The typing and semantics for these lenses is provided in Figure~\ref{fig:dnf-lens-alternate-alternate-semantics}.
DNF Lenses express exactly the same amount as standard lenses.
\begin{restatable}[Completeness of DNF Lenses]{theorem}{dnflc}
\label{thm:completeness-dnf-lenses}
If there exists a derivation of $\FullContext \DerivesLens \Lens \OfType \MapsBetweenTypeOf{\Regex}{\RegexAlt} \HasSemantics \PutRight,\PutLeft$,
then there exists a derivation of $\FullContext \DerivesDNFLens \DNFLens \OfType \MapsBetweenTypeOf{\DNFRegex}{\DNFRegexAlt} \HasSemantics \PutRight,\PutLeft$ such that
$\LanguageOf{\RegexContext}{\DNFRegex}=\LanguageOf{\RegexContext}{\Regex}$, and
$\LanguageOf{\RegexContext}{\DNFRegexAlt}=\LanguageOf{\RegexContext}{\RegexAlt}$.
\end{restatable}

\begin{restatable}[Soundness of DNF Lenses]{theorem}{dnfls}
\label{thm:soundness-dnf-lenses}
If there exists a derivation of $\FullContext \DerivesDNFLens \DNFLens \OfType \MapsBetweenTypeOf{\DNFRegex}{\DNFRegexAlt} \HasSemantics \PutRight,\PutLeft$,
then there exists a derivation of $\FullContext \DerivesLens \Lens \OfType \MapsBetweenTypeOf{\Regex}{\RegexAlt} \HasSemantics \PutRight,\PutLeft$ such that
$\LanguageOf{\RegexContext}{\Regex}=\LanguageOf{\RegexContext}{\DNFRegex}$, and
$\LanguageOf{\RegexContext}{\RegexAlt}=\LanguageOf{\RegexContext}{\DNFRegexAlt}$.
\end{restatable}

DNF lenses handle the equivalences of concatenation associativity,
or associativity, or commutativity, and and-or distributivity.
Furthermore, they do not require composition to be complete,
though we do allow for concatenation of lenses in the lens context.
While these are not needed for full expressibility because of Theorem~\ref{thm:empty-context-regex-completeness}
and Theorem~\ref{thm:completeness-dnf-lenses},
we choose to use these contextual lenses and regular expression.
This is under the assumption that if lenses are present
in the context, they are likely to be used.
However, this does add in additional choices for synthesis, but we've found
that in practice, it simplifies the algorithm (TODO: this analysis).
Because of this, the choices needed to be done by the synthesis algorithm
are fairly minimal.  There are only
a few types of equivalences related to stars and use of regular expressions
in the context that need to be explored for completeness.
Furthermore, there is the difficulty of determining which permutations create
well typed regular expressions, but our implentation reduces the problem of
finding a valid permutation to ordering the terms under a specific order.

\input{fig_regex-equivalence-rules}
Two regular expressions are said to be equivalent if they are equivalent through
any of the rules specified in Figure~\ref{fig:regex-equivalence-rules}.
Indeed, after ordering we have put our regular expressions in a very normal form.
If we normalize two regular expressions, and those two regular expressions are
equivalent through any number of applications of equivalences 1 through 8, then
their ordered DNF regular expressions will be equivalent.
What's more, we can use the ordering on ordered DNF regular expressions to create another
equivalence class.  The ordering on DNF regular expressions says two ordered
dnf regular expressions are equal if, and only if, there exists a lens between
then, where the lens typing doesn't include the DNF Rewrite Lens rule.  The purpose
of the DNF Rewrite rule is to handle the other types of equivalences.
